\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-07-09 paper.tex
IROS review: extend discussion on knowledge model

Relying on RDF triples and Description Logics has major advantages like the
availability of numerous open-source libraries to manipulate the ontology,
interoperability with several major on-line knowledge bases (like {\sc
OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
guarantee of decidability (it is always possible to classify a Description
Logics ontology).

It has also notable limitations: RDF triples imply only binary predicates
(\stmt{subject predicate object}), which constrains the expressiveness of the
system or leads to cumbersome reifications. Alternative exists (like {\sc
KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
languages like {\sc Prolog}, at the price, however, of other limitations, like
closed-world reasoning or immutable T-Box. The classification
performance is another issue: from our experience, with an ontology sized for a
standard experiment (about 100 classes and 200 instances), classification
typically take about 100ms, which becomes problematic during interactions.
Besides, the performances are difficult to predict, since a seemingly
inoffensive new statement may indirectly change radically the logical
complexity of the whole knowledge model and lead to notable degradation of
classification time.

This knowledge model also largely exclude representation of continuous
phenomena (like time) or uncertain phenomena. When required, these are managed
inside the corresponding components, and are not exposed at the system level.

Amongst the features of {\sc Oro}, the system can dynamically create
new independent knowledge models (\ie independent instantiations of the
ontology). This feature is used for instance to store separately the beliefs of
each of the agents the robot interacts with (see section~\ref{sect|tom}).

>>>>

Relying on RDF triples and Description Logics has major advantages like the
availability of numerous mature open-source libraries to manipulate the ontology,
interoperability with several major on-line knowledge bases (like {\sc
OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
guarantee of decidability (it is always possible to classify a Description
Logics ontology).

It has also notable limitations, both fundamental (the suitability
of Description Logics when reasoning on --typically non-monotonic-- commonsense
knowledge is questionable) and practical: RDF triples imply only binary predicates
(\stmt{subject predicate object}), which constrains the expressiveness of the
system or leads to cumbersome reifications. Alternative exists (like {\sc
KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
languages like {\sc Prolog}, at the price, however, of other limitations, like
closed-world reasoning or immutable T-Box. The classification
performance is another issue: from our experience, with an ontology sized for a
standard experiment (about 100 classes and 200 instances), classification
typically take about 100ms, which becomes problematic during interactions.
Besides, the performances are difficult to predict, since a seemingly
inoffensive new statement may indirectly change radically the logical
complexity of the whole knowledge model and lead to notable degradation of
classification time.

This knowledge model also largely exclude representation of continuous
phenomena (like time) or uncertain phenomena. When required (for instance for
action recognition), these are managed inside the corresponding components, and
are not exposed at the semantic level.

Tools to manipulate and reason over ontologies are readily available, mature,
and already well accepted in the robotic community~\cite{Tenorth2009a,
Lim2011}. While alternatives like \emph{Answer Set Programming} have also been
successfully investigated in robotics~\cite{Chen2010,Erdem2012}, in particular
to deal with non-monotonic reasoning, we did not actually hit any brick wall
while working with OWL ontologies. We may reconsider this choice at a later
stage, but until now it has proven an effective framework to quickly explore
implementations of new cognitive abilities (for instance, it has been
conceptually and technically easy to add support for independent knowledge
models, one per the agents the robot interacts with --see
section~\ref{sect|tom}).

Besides, because ontologies and RDF statements are relatively simple concepts
to grasp, it also effectively helped to grow awareness amongst colleagues on
the significance of the ``semantic level'' when developing new components for
the robot.
